---
title: "Loops and Functions"
author: "Alex Vand"
date: "8/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Part 1: For loops revisited

Create two sequences, one called `weekdays` that contains days of the week (“Monday”, “Tuesday”, “Wednesday”, etc.) and one called `transects` that contains the series of transect names “Transect A”, “Transect B,”Transect C". Write a nested for loop that creates a matrix containing the following:


```{r}
weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")
transects <- c("Transect A", "Transect B", "Transect C")
matrix <- matrix(nrow = length(weekdays), ncol = length(transects))

for(i in seq_along(weekdays)) {
  for(j in seq_along(transects)) {
    matrix[i, j] <- paste(weekdays[i], "-", transects[j])
  }
}
```


## Part 2: Functions introduced

### Task 2
Write a function called `force` that calculates a force (in Newtons), given inputs of mass (in kg) and acceleration (in $ms^2$ (recall: $F=ma$)), and returns a statement `“The resulting force is ___ Newtons.”`

```{r}
force <- function(mass, acceleration) {
  force = mass * acceleration
  print(paste("The resulting force is", force, "Newtons."))
}

force(40, 6)
```

### Task 3
Write a function called `name_column` that, given a data frame name and a number, will return the column name associated with that column number. Add an error message if the column number is < 1, or if the number exceeds the number of columns in the data frame. Test the function using the `mtcars` data frame.

```{r}
name_column <- function(df, col_num) {
  return(colnames(df[col_num]))
  
  if(col_num < 1 | col_num > lenth(df)) {
    stop("Incorrect column number")
  }
}
name_column(mtcars, 8)
```

### Task 4

```{r}
sci_name <- c("Chanos chanos", "Sphyraena barracuda", "Caranx ignobilis")
common_name <- c("Milkfish", "Giant barracuda", "Giant trevally")
a_est <- c(0.0905, 0.0181, 0.0353)
b_est <- c(2.52, 3.27, 3.05)

fish_parms <- data.frame(sci_name, common_name, a_est, b_est)

fish_weight <- function(fish_name, tot_length) {
  select_fish <- dplyr::filter(fish_parms, common_name == fish_name)
  a = select_fish$a_est
  b = select_fish$b_est
  weight <- a * tot_length ^ b
  return(weight)
}

fish_weight("Milkfish", 6)
fish_weight("Giant barracuda", 20)

lengths = seq(from = 0, to = 100, by = 1)

milkfish_weights <- fish_weight("Milkfish", lengths)
milkfish_weights
```

### Task 5

```{r}
wave_power_deep <- function(height, period) {
  P_deep <- 0.5 * height ^ 2  * period
  return(P_deep)
}

wave_height <- seq(from = 0, to = 3, by = 0.2)

wave_power_deep(wave_height, 8)
```

### Task 6

```{r}
wave_power <- function(depth, height, period) {
  if(depth > 12) {
      P_deep <- 0.5 * height ^ 2 * period
      message("Deep wave equation used.")
      return(P_deep)
}
  
  else if(depth <= 12) {
      P_shallow <- 0.81 * height ^ 2 * period
      message("Shallow wave equation used.")
      return(P_shallow)
      
}
}

wave_power(5, 12, 12)
```







